---
title: "PAN-E Return to nature"
author: "Cerren Richards"
date: "14/05/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we use the Google mobility data to find the t-value for all countries before and after COVID-19 confinements.

### Google Data

**Data Description**:

* How visits and length of stay at different places change compared to a baseline.
    - Parks: national parks, public beaches, marinas, dog parks, plazas, and public gardens.
    - Residential: places of residence
    
**Baseline**:

* Median value, for the corresponding day of the week, during the 5- week period Jan 3–Feb 6, 2020

**Download data and methods**:

* https://www.google.com/covid19/mobility/
* https://support.google.com/covid19-mobility/answer/9824897?hl=en&ref_topic=9822927


## Global trend - parks, residential and transport

```{r, error=FALSE, warning=FALSE}
google <- read.csv("Global_Mobility_Report (2).csv")

# R reads Namibia's iso2 code is as an "is.na" object, so we will rename it to NA
google$country_region_code[is.na(google$country_region_code)] <-"NA"

# Add the country codes so we ensure everything matches 
library(countrycode)
google$ISO2 <- google$country_region_code
google$ISO3 <- countrycode(google$ISO2, origin = 'iso2c', destination = 'iso3c')
google$country <- countrycode(google$ISO2, origin = 'iso2c', destination = 'country.name')

# Add Julian date
google$julian<- (strptime(google$date, "%Y-%m-%d")$yday) + 1 # need to add 1 day because 2020 is a leap year

library(naniar)
# remove the subregions and only keep the overall country pattern
google<- google %>% replace_with_na(replace = list(sub_region_1 = ""))
google<- google[is.na(google$sub_region_1),]

## Calculate global median movement at parks, transit and residential
google_median <- google %>% group_by(date) %>% 
                    summarise(`Parks & Beaches`= 
                                median(parks_percent_change_from_baseline, na.rm = TRUE),
                              `Grocery & Pharmacy` = 
                                median(grocery_and_pharmacy_percent_change_from_baseline, na.rm = TRUE),
                              Residential = 
                                median(residential_percent_change_from_baseline, na.rm = TRUE))

## Define "date" as a date 
google_median$date <- as.Date(google_median$date)

library(tidyr)
## Rearrange for plotting
google_median <- google_median %>% gather(type, change, `Parks & Beaches`:Residential)

## Calculate country median movement at parks, transit and residential
#google_countries <- google %>% group_by(date, country_region) %>% 
#                    summarise(`Parks & Beaches`= median(parks_percent_change_from_baseline, na.rm = TRUE),
#                              `Transit Stations` = median(transit_stations_percent_change_from_baseline, na.rm = TRUE),
#                              Residential = median(residential_percent_change_from_baseline, na.rm = TRUE))


google_countries <- google %>% select(date,country_region,ISO3, 
                                      `Parks & Beaches`= 
                                        parks_percent_change_from_baseline,
                                      `Grocery & Pharmacy` = 
                                        grocery_and_pharmacy_percent_change_from_baseline,
                                          Residential = 
                                        residential_percent_change_from_baseline) 



library(ggplot2);library(viridis); library(ggpubr);library(sf);library("rnaturalearth"); library("rnaturalearthdata")
# extract continent info
# data for the world
world <- ne_countries(scale = "medium", returnclass = "sf")
world <- world %>% rename(ISO3 = iso_a3) 
region_un <- world$region_un
continent <- world$continent
ISO3 <- world$ISO3
continent<- tibble(ISO3, region_un, continent)
# join to circle data
google_countries <- left_join(google_countries, continent, by = "ISO3")
# assign Tuvalu to Oceania
google_countries$region_un[is.na(google_countries$region_un)] <-"Oceania"

## Define "date" as a date 
google_countries$date <- as.Date(google_countries$date)

## Rearrange for plotting
google_countries <- google_countries %>% gather(type, change, `Parks & Beaches`:Residential)



  
ggplot(google_median, aes(date, `Parks & Beaches`)) +
  geom_line(data = google_countries, aes(group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  labs(x = "", 
       y = "Change in length of visit (%)",
       title = "Parks & Beaches")

ggplot(google_median, aes(date, `Grocery & Pharmacy`)) +
  geom_line(data = google_countries, aes(group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  labs(x = "", 
       y = "Change in length of visit (%)",
       title = "Grocery & Pharmacy")

ggplot(google_median, aes(date, Residential)) +
  geom_line(data = google_countries, aes(group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  labs(x = "", 
       y = "Change in length of visit (%)",
       title = "Residential")



ggplot() +
  geom_line(data = google_countries, aes(date, `Parks & Beaches`, group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  facet_wrap(~region_un, scales = "free", ncol = 3)+
  labs(x = "", 
       y = "Change in length of visit (%)",
       title = "Parks & Beaches")+
  theme(axis.text.x = element_text(angle = 90))


ggplot() +
  geom_line(data = google_countries, aes(date, `Grocery & Pharmacy`, group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  facet_wrap(~region_un, scales = "free", ncol = 3)+
  labs(x = "", 
       y = "Change in length of visit (%)",
       title = "Grocery & Pharmacy")+
  theme(axis.text.x = element_text(angle = 90))

ggplot() +
  geom_line(data = google_countries, aes(date, Residential, group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  facet_wrap(~region_un, scales = "free", ncol = 3)+
  labs(x = "", 
       y = "Change in length of visit (%)",
       title = "Residential")+
  theme(axis.text.x = element_text(angle = 90))


ggplot(google_median, aes(date, change)) +
  geom_line(data = google_countries, aes(group = country_region), colour = alpha("grey", 0.5))+
  geom_line(size = 1)+
  geom_hline(yintercept = 0, linetype = "dashed")+
    scale_x_date(date_breaks = "4 week", date_labels = "%b %d")+
      theme_bw(base_size = 15)+ # set the background theme                                              
  theme(panel.grid.major = element_blank(), # remove the major lines
        panel.grid.minor = element_blank())+ # remove the minor lines
  facet_wrap(~type, scales = "free", nrow = 3)+
  labs(x = "", 
       y = "Change in length of visit (%)")+
    theme(strip.text.x = element_text(size = 15, color = "white"), 
          strip.background = element_rect(fill="black"))

ggsave("Parks_grocety_residential.pdf", 
       dpi = 600, 
       width = 160, height = 200, unit = "mm")

      # subtitle = "Black line = median across 135 countries. Grey lines = countries.")


```



